A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-11-01 linear.tex
adding kz=0 special case in response to inutsuka-san's query

\section{Compressible streaming instability}\label{si}
In the special case of an unstratified disk (neglecting the vertical
component of the stellar gravity), we may also Fourier analyze in $z$
to obtain an algebraic dispersion relation of the form  
$\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
vertical wavenumber. The coefficients $c_j$ can be read 
off Eq. \ref{streaming_dispersion} in  Appendix \ref{compressible_streaming}. 
There we also show that this dispersion relation reduces to that for
the streaming instability (SI) in the limit of incompressible gas and small
$\tstop$ \citep{youdin05a,jacquet11}.

{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
growth rates have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.

>>>>

\section{Dust-drag instabilities}\label{si}
In the special case of an unstratified disk (neglecting the vertical
component of the stellar gravity), we may also Fourier analyze in $z$
to obtain an algebraic dispersion relation of the form  
$\sum_{j=0}^{5}c_j(k_x,k_z)\sigma^j = 0$, where $k_z$ is a real
vertical wavenumber. The coefficients $c_j$ can be read 
off Eq. \ref{streaming_dispersion} in  Appendix \ref{compressible_streaming}. 
There we also show that this dispersion relation reduces to that for
the streaming instability (SI) in the limit of incompressible gas and small
$\tstop$ \citep{youdin05a,jacquet11}.

\subsection{Overstable dusty epicycles}
In the special case of $k_z=0$, the incompressible dispersion
relation, Eq. \ref{streaming_incompressible} reduces to 
\begin{align}      %streaming_incompressible
\ii \sigma^2 - \frac{\sigma}{\tepsilon \taus}  - \left(\ii \kappa^2 +
  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
\end{align}
If the quadratic term is neglected, then $\imag{\sigma}<0$, implying
stability, as concluded by \cite{youdin05a}.

However, retaining all the terms (but still in the incompressible
limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
pure oscillations with $\sigma = -\omega$ provided that 
\begin{align*}
  \omega &= \pm \kappa,\\
  \taus & = \frac{\epsilon\kappa}{\tepsilon k _x\left|F_r\right|}
         \equiv \tau_{\mathrm{s}*}.
\end{align*}
If we linearize about this state of marginal stability by writing
$\sigma = - \kappa + \ii \dd s$ and $\taus = \tau_{\mathrm{s}*} +
\dd\taus$ and assume $|\dd s|\ll \kappa$ then 
\begin{align*} 
\dd s = \frac{2\tepsilon }{1 +
  4\tepsilon^2\tau_{\mathrm{s}*}^2}\dd \taus. 
\end{align*}
Thus given $\tepsilon$ and $k_x$, there are overstable modes with
stopping times slightly larger than $\tau_{\mathrm{s}*}$.

%{\bf note: several approx used}

\subsection{Streaming instability}

{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
We solve Eq. \ref{streaming_dispersion} for selected cases where SI  
growth rates have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_{x,z} = \eta r k_{x,z}$ where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.
